---
title: "R Notebook Wrangle"
output: html_notebook
---

## Data Cleaning with R

The process of identifying, correcting, or removing inaccurate raw data for downstream purposes. Or, more colloquially, an unglamorous yet wholely necessary first step towards an analysis-ready.

### Check List

-   [ ] Familiarize yourself with the data set

-   [ ] Check for structural errors

-   [ ] Check for data irregularities

-   [ ] Decide how to deal with missing values

-   [ ] Document data versions and changes made

### Step 1: Familiarize yourself with the data set

An important "pre-data cleaning" Step is **Domain Knowledge**.

> Be sure to read up or ask an expert **before** diving into data cleaning if your variables don't make sense to you at first.sure to read up or ask an expert **before** diving into data cleaning if your variables don't make sense to you at first.

```{r}
file.info("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")$size
```

The data file we're utilizing is 1104203 bytes or 1.01MB.

```{r}
df = read.csv("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")

```

```{r}
head(df)
```

```{r}
str(df) # Show structure of data frame
```

#### Simple Fast Exploratory Data Analysis in R with DataExplorer Package

```{r}
#Install if the package doesn't exist 
update.packages()

```

```{r}
install.packages('DataExplorer')

```

```{r}
if (!require(devtools)) install.packages("devtools")
```

```{r}
library(devtools)
```

```{r}
library("DataExplorer")
```

# Why Do we need clean data?

-   [ ] Access Data
-   [ ] Explore and Process Data
-   [ ] Extract Insights
-   [ ] Report Insights

## Data type constraints

-   Text or in R is character
-   Integer - integer - numeric
-   Binary - logical
-   Category - factor
-   Date - date

```{r}
edidiv <- read.csv("./CC-RBasics/edidiv.csv")
head(edidiv)
```

### Glimpsin at data tupes

```{r}
library(dplyr)
glimpse(edidiv)
```

**Note** : This help you to see the data type in the data set.

### Checking data types

-   `is.character()`
-   `is.numeric()`
-   `is.logical`
-   for all `data types`

```{r}
is.numeric(edidiv$year)
```

```{r}
install.packages("assertive")
```

```{r}

library(assertive)
assert_is_factor(edidiv$taxonName)
```

```{r}
class(edidiv$organisationName)
```

```{r}
mean(edidiv$taxonName)
```

### Comma problems

```{r}
library("tidyverse")
```

```{r}
comma_problems <- c("5,454", "5,456", "3,456", "6,789")

```

```{r}
library(stringr)
```

```{r}
revenue_trimed = str_remove(comma_problems, ",")
revenue_trimed
```

### Character to Numeric

```{r}
as.numeric(revenue_trimed)
```

### Putting it together

```{r}
edidiv %>% 
  mutate( 
    grid_mod = as.numeric(
      str_remove(gridReference, "NT")
    )
  ) 
```

# Google class - data cleaning

The next packages are important for data cleaning process in R:

```{r}
install.packages("here")
```

```{r}
library('here')
```

The goal of the here package is to enable easy file referencing in project-oriented workflows.\
In contrast to using setwd(), which is fragile and dependent on the way you organize your files,\
here uses the top-level directory of a project to easily build paths to files.

```{r}
install.packages("skimr")
```

```{r}
library("skimr")
```

skimr is designed to provide summary statistics about variables in data frames, tibbles, data tables and vectors. It is opinionated in its defaults, but easy to modify.

In base R, the most similar functions are summary() for vectors and data frames and fivenum() for numeric vectors:

```{r}
install.packages("janitor")
```

```{r}
library("janitor")
```

The janitor package is a R package that has simple functions for examining and cleaning dirty data. It was built with beginning and intermediate R users in mind and is optimised for user-friendliness

```{r}
install.packages("dplyr")
```

```{r}
library("dplyr")
```

**dplyr** is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges:

-   mutate() adds new variables that are functions of existing variables
-   select() picks variables based on their names.
-   filter() picks cases based on their values.
-   summarise() reduces multiple values down to a single summary.
-   arrange() changes the ordering of the rows.

```{r}
install.packages("palmerpenguins")
```

```{r}
install.packages("remotes")
```

```{r}
library("remotes")
```

\`
