---
title: "R Notebook"
output: html_notebook
---

# Visualizations in R

### Popular graph packages:

-   ggplot2

-   Plotly

-   Lattice

-   RGL

-   Dygraphs

-   Leaflet

-   Highcharter

-   Patchwork

-   gganimate

-   ggridges

### Benefits of ggplot2:

-   Create different types of plots

-   Customize the look and feel of plots

-   Create high quality visuals

-   Combine data manipulation and visualization

### Concepts in ggplot2:

-   **Aesthetics** --\> A visual property of an object in your plot
    (shape, color)

-   **Geoms** ---\> The geometric object used to represent your data
    (lines, bars)

-   **Facets** ---\> Let you display smaller groups, or subsets, of your
    data

-   **Labels and annotation**s ---\> Let you Customize your plot

## Create graph basic

```{r}
install.packages("ggplot2")

install.packages("palmerpenguins")
```

```{r}
library(ggplot2)
library(palmerpenguins)
```

```{r}
data(penguins)
```

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))
```

-   geom_point() --\> Scatterplots graph

-   geom_bar() --\> Bars graph

Each geom function in ggplot2 takes a **mapping argument**. This defines
how variables in your dataset are mapped to visual properties. The
mapping argument is always paired with the **aes() function**.

# Getting started with ggplot()

```{r}
library("palmerpenguins")
library("ggplot2")
```

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
body <- na.omit(penguins$body_mass_g)
flipper <- na.omit(penguins$flipper_length_mm)

cor(flipper, body)
```

( + ) is used for add layer in the graph.

1.  Start with the ggplot function and choose a dataset to work with
2.  Add a geom_function to display your data
3.  Map the variables you want to plot in the arguments of the aes()
    function

```{r}
# Help is very useful 
?geom_bar
```

```{r}
# geom_bar is designed to make it easy to create bar charts that show
# counts (or sums of weights)
g <- ggplot(mpg, aes(class))
# Number of cars in each class:
g + geom_bar()
# Total engine displacement of each class
g + geom_bar(aes(weight = displ))
# Map class to y instead to flip the orientation
ggplot(mpg) + geom_bar(aes(y = class))
```

# Enhancing visualizations in R

#### Colors

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color = species))
```

#### Shape

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color= species, 
                         shape = species))
```

#### Size

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color= species, 
                         shape = species, size = species))
```

```{r}

```

#### Alpha

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, shape = species, 
                         alpha = species))
```

#### Colors all points

```{r}
# Note color is out aes function
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, 
                         shape = species, size = species), color = "blue")
```

**Ggplot2** is an R package that allows you to create different types of
data visualizations right in your R workspace. In ggplot2, an
**aesthetic** is defined as a visual property of an object in your
plot. 

There are three aesthetic attributes in ggplot2:

-   **Color**: this allows you to change the color of all of the points
    on your plot, or the color of each data group

-   **Size**: this allows you to change the size of the points on your
    plot by data group

-   **Shape**: this allows you to change the shape of the points on your
    plot by data group

# Doing more with ggplot

```{r}
ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
ggplot(data = penguins) +
  geom_smooth(mapping=aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
ggplot(data = penguins) +
  geom_smooth(mapping=aes(x=flipper_length_mm, y=body_mass_g, linetype =species )) 
  

ggplot(data = penguins) +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
ggplot(data = penguins) +
  geom_jitter(mapping=aes(x=flipper_length_mm, y=body_mass_g))
```

```{r}
ggplot(data = diamonds ) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = diamonds ) + 
  geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds ) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds ) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

### Smoothing

```{r}
ggplot(penguins, aes(x=flipper_length_mm, 
y= body_mass_g)) +
    geom_point() +
    geom_smooth()
```

**Smoothing** enables the detection of a data trend even when you can't
easily notice a trend from the plotted data points. Ggplot2's smoothing
functionality is helpful because it adds a **smoothing line** as another
layer to a plot; the smoothing line helps the data to make sense to a
casual observer.

| **Type of smoothing**             | **Description**                                   | **Example code** |
|-----------------------------------|---------------------------------------------------|------------------|
| **Loess smoothing**               | The loess smoothing process is best for smoothing |                  |
| plots with less than 1000 points. | ggplot(data, aes(x=, y=))+                        |                  |

                                                                                 geom_point() +               
                                                                                 geom_smooth(method="loess")  |

| **Gam smoothing** \| Gam smoothing, or generalized additive model smoothing, is useful for smoothing plots with a large number of points.  \| ggplot(data, aes(x=, y=)) + geom_point() +         geom_smooth(method="gam", formula = y \~s(x)) \|

```{r}
ggplot(penguins, aes(x=flipper_length_mm, 
y= body_mass_g))+ 
  geom_point() +       
  geom_smooth(method="loess")
```

```{r}
ggplot(penguins, aes(x=flipper_length_mm, 
y= body_mass_g)) + 
  geom_point() +        
  geom_smooth(method="gam", 
formula = y ~s(x))
```

# Aesthetics and facets

-   facet_wrap()

-   facet_grid()

### Face wrap

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(
    x = flipper_length_mm, y= body_mass_g, color = species )
  ) + 
  facet_wrap(~species)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(
    x = color, fill= cut )
  ) + 
  facet_wrap(~cut)
```

### Face grip

Two variables

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(
    x = flipper_length_mm, y= body_mass_g, color = species )
  ) + 
  facet_grid(sex~species)
```

## E​xample of filtering data for plotting

Filtering your data before you plot it allows you to focus on specific
subsets of your data and gain more targeted insights. To do this, just
include the dplyr filter() function in your ggplot syntax. 

```{r}
library(nycflights13)
```

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
#> `summarise()` ungrouping output (override with `.groups` argument)
delay <- filter(delay, count > 20, dest != "HNL")

# It looks like delays increase with distance up to ~750 miles 
# and then decrease. Maybe as flights get longer there's more 
# ability to make up delays in the air?
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```

# Annotation layer

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point()
p + annotate("text", x = 4, y = 25, label = "Annotate some text here!")
```

```{r}
p + annotate("text", x = 2:5, y = 25, label = "Some text")
```

```{r}
p + annotate("rect", xmin = 3, xmax = 4.2, ymin = 12, ymax = 21,
  alpha = .2)
```

```{r}
p + annotate("segment", x = 2.5, xend = 4, y = 15, yend = 25,
  colour = "blue")
```

```{r}
p + annotate("pointrange", x = 3.5, y = 20, ymin = 12, ymax = 28,
  colour = "red", size = 1.5)
```

```{r}
p + annotate("text", x = 4, y = 25, label = "italic(R) ^ 2 == 0.75",
  parse = TRUE)
```

```{r}
# basic graph
p <- ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point()
 
# a data frame with all the annotation info
annotation <- data.frame(
   x = c(2,4.5),
   y = c(20,25),
   label = c("label 1", "label 2")
)

# Add text
p + geom_text(data=annotation, aes( x=x, y=y, label=label),                 , 
           color="orange", 
           size=7 , angle=45, fontface="bold" )

# Note: possible to shorten with annotate:
# p +
#  annotate("text", x = c(2,4.5), y = c(20,25), 
#           label = c("label 1", "label 2") , color="orange", 
#           size=7 , angle=45, fontface="bold")

# Right chart: using labels
p + geom_label(data=annotation, aes( x=x, y=y, label=label),                 , 
           color="orange", 
           size=7 , angle=45, fontface="bold" )
```

```{r}
# Add rectangles
p + annotate("rect", xmin=c(2,4), xmax=c(3,5), ymin=c(20,10) , ymax=c(30,20), alpha=0.2, color="blue", fill="blue")

# Add segments
p + annotate("segment", x = 1, xend = 3, y = 25, yend = 15, colour = "purple", size=3, alpha=0.6)

# Add arrow
p + annotate("segment", x = 2, xend = 4, y = 15, yend = 25, colour = "pink", size=3, alpha=0.6, arrow=arrow())
```

```{r}
p + 
  # horizontal
  geom_hline(yintercept=25, color="orange", size=1) + 
  # vertical
  geom_vline(xintercept=3, color="orange", size=1)
```

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = factor(cyl))) + 
  labs(
    x = "Engine displacement (litres)", 
    y = "Highway miles per gallon", 
    colour = "Number of cylinders",
    title = "Mileage by engine size and cylinders",
    subtitle = "Source: http://fueleconomy.gov"
  )
```

```{r}
install.packages("ggtext")
```

```{r}
library("ggtext")
```

```{r}
df <- data.frame(x = 1:3, y = 1:3)
base <- ggplot(df, aes(x, y)) + 
  geom_point() + 
  labs(x = "Axis title with *italics* and **boldface**")

base 
base + theme(axis.title.x = ggtext::element_markdown())
```

```{r}
df <- data.frame(
  x = c(1, 1, 2, 2, 1.5),
  y = c(1, 2, 1, 2, 1.5),
  text = c(
    "bottom-left", "bottom-right", 
    "top-left", "top-right", "center"
  )
)
ggplot(df, aes(x, y)) +
  geom_text(aes(label = text))
ggplot(df, aes(x, y)) +
  geom_text(aes(label = text), vjust = "inward", hjust = "inward")
```

```{r}
df <- data.frame(trt = c("a", "b", "c"), resp = c(1.2, 3.4, 2.5))
ggplot(df, aes(resp, trt)) + 
  geom_point() + 
  geom_text(aes(label = paste0("(", resp, ")")), nudge_y = -0.25) + 
  xlim(1, 3.6)
```

```{r}
ggplot(economics, aes(date, unemploy)) + 
  geom_line()
```

```{r}
presidential <- subset(presidential, start > economics$date[1])

ggplot(economics) + 
  geom_rect(
    aes(xmin = start, xmax = end, fill = party), 
    ymin = -Inf, ymax = Inf, alpha = 0.2, 
    data = presidential
  ) + 
  geom_vline(
    aes(xintercept = as.numeric(start)), 
    data = presidential,
    colour = "grey50", alpha = 0.5
  ) + 
  geom_text(
    aes(x = start, y = 2500, label = name), 
    data = presidential, 
    size = 3, vjust = 0, hjust = 0, nudge_x = 50
  ) + 
  geom_line(aes(date, unemploy)) + 
  scale_fill_manual(values = c("blue", "red")) +
  xlab("date") + 
  ylab("unemployment")
```

```{r}
install.packages("directlabels")
```

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()

ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = class), method = "smart.grid")
```

```{r}
install.packages("ggforce")
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() + 
  ggforce::geom_mark_ellipse(aes(label = cyl, group = cyl))
```

# Saving your visualizations

```{r}
ggsave("datavizsave.png")
```

| **Example of using png()**                        | **Example of using pdf()**                            |
|---------------------------------------------------|-------------------------------------------------------|
| png(file = "exampleplot.png", bg = "transparent") 
 plot(1:10)                                         
 rect(1, 5, 3, 7, col = "white")                    
 dev.off()                                          | pdf(file = "/Users/username/Desktop/example.pdf",     
                                                     width = 4,                                             
                                                     height = 4)                                            
                                                     plot(x = 1:10,                                         
                                                      y = 1:10)                                             
                                                     abline(v = 0)                                          
                                                     text(x = 0, y = 1, labels = "Random text")             
                                                     dev.off()                                              |

```{r}
if (TRUE) {
ggplot(mtcars, aes(mpg, wt)) + geom_point()

ggsave("mtcars.pdf")
ggsave("mtcars.png")

ggsave("mtcars.pdf", width = 4, height = 4)
ggsave("mtcars.pdf", width = 20, height = 20, units = "cm")

# delete files with base::unlink()
unlink("mtcars.pdf")
unlink("mtcars.png")

# specify device when saving to a file with unknown extension
# (for example a server supplied temporary file)
file <- tempfile()
ggsave(file, device = "pdf")
unlink(file)

# save plot to file without using ggsave
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
png("mtcars.png")
print(p)
dev.off()

}
```
