---
title: "R Notebook Wrangle"
output: html_notebook
---

## Data Cleaning with R

The process of identifying, correcting, or removing inaccurate raw data
for downstream purposes. Or, more colloquially, an unglamorous yet
wholely necessary first step towards an analysis-ready.

### Check List

-   [ ] Familiarize yourself with the data set

-   [ ] Check for structural errors

-   [ ] Check for data irregularities

-   [ ] Decide how to deal with missing values

-   [ ] Document data versions and changes made

### Step 1: Familiarize yourself with the data set

An important "pre-data cleaning" Step is **Domain Knowledge**.

> Be sure to read up or ask an expert **before** diving into data
> cleaning if your variables don't make sense to you at first.sure to
> read up or ask an expert **before** diving into data cleaning if your
> variables don't make sense to you at first.

```{r}
file.info("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")$size
```

The data file we're utilizing is 1104203 bytes or 1.01MB.

```{r}
df = read.csv("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")

```

```{r}
head(df)
```

```{r}
str(df) # Show structure of data frame
```

```{r}
#Install if the package doesn't exist 
update.packages()

```

# Why Do we need clean data?

-   [ ] Access Data
-   [ ] Explore and Process Data
-   [ ] Extract Insights
-   [ ] Report Insights

## Data type constraints

-   Text or in R is character
-   Integer - integer - numeric
-   Binary - logical
-   Category - factor
-   Date - date

```{r}
edidiv <- read.csv("./CC-RBasics/edidiv.csv")
head(edidiv)
```

### Glimpsin at data tupes

```{r}
library(dplyr)
glimpse(edidiv)
```

**Note** : This help you to see the data type in the data set.

### Checking data types

-   `is.character()`
-   `is.numeric()`
-   `is.logical`
-   for all `data types`

```{r}
is.numeric(edidiv$year)
```

```{r}
install.packages("assertive")
```

```{r}

library(assertive)
assert_is_factor(edidiv$taxonName)
```

```{r}
class(edidiv$organisationName)
```

```{r}
mean(edidiv$taxonName)
```

### Comma problems

```{r}
library("tidyverse")
```

```{r}
comma_problems <- c("5,454", "5,456", "3,456", "6,789")

```

```{r}
library(stringr)
```

```{r}
revenue_trimed = str_remove(comma_problems, ",")
revenue_trimed
```

### Character to Numeric

```{r}
as.numeric(revenue_trimed)
```

### Putting it together

```{r}
edidiv %>% 
  mutate( 
    grid_mod = as.numeric(
      str_remove(gridReference, "NT")
    )
  ) 
```

# Google class - data cleaning

The next packages are important for data cleaning process in R:

```{r}
install.packages("here")
```

```{r}
library('here')
```

The goal of the here package is to enable easy file referencing in
project-oriented workflows.\
In contrast to using setwd(), which is fragile and dependent on the way
you organize your files,\
here uses the top-level directory of a project to easily build paths to
files.

```{r}
install.packages("skimr")
```

```{r}
library("skimr")
```

skimr is designed to provide summary statistics about variables in data
frames, tibbles, data tables and vectors. It is opinionated in its
defaults, but easy to modify.

In base R, the most similar functions are summary() for vectors and data
frames and fivenum() for numeric vectors:

```{r}
install.packages("janitor")
```

```{r}
library("janitor")
```

The janitor package is a R package that has simple functions for
examining and cleaning dirty data. It was built with beginning and
intermediate R users in mind and is optimised for user-friendliness

```{r}
install.packages("dplyr")
```

```{r}
library("dplyr")
```

**dplyr** is a grammar of data manipulation, providing a consistent set
of verbs that help you solve the most common data manipulation
challenges:

-   mutate() adds new variables that are functions of existing variables
-   select() picks variables based on their names.
-   filter() picks cases based on their values.
-   summarise() reduces multiple values down to a single summary.
-   arrange() changes the ordering of the rows.

```{r}
install.packages("palmerpenguins")
```

```{r}
install.packages("remotes")
```

```{r}
library("remotes")
```

```{r}
library("palmerpenguins")
```

### Data explored

```{r}
skim_without_charts(penguins)
```

```{r}
glimpse(penguins)
```

```{r}
head(penguins)
```

```{r}
penguins %>%
  select(species)
```

```{r}
penguins %>% 
  select(-species) # Excluded species from penguins dataset

```

### Transform

```{r}
penguins_fail <- rename( penguins, isl = island) # Rename columns
```

```{r}
rename_with(penguins, toupper) # Change toupper and tolower
```

## File Convention

#### **Do**

-   Keep your filenames to a reasonable length

-   Use underscores and hyphens for readability

-   Start or end your filename with a letter or number

-   Use a standard date format when applicable; example: YYYY-MM-DD

-   Use filenames for related files that work well with default
    ordering; example: in chronological order, or logical order using
    numbers first

| E​xamples of good filenames     |
|---------------------------------|
| 2020-04-10_march-attendance.R   |
| 2021_03_20_new_customer_ids.csv |
| 01_data-sales.html              |
| 02_data-sales.html              |

### **Arithmetic operators**

| Operator | Description                                                 | Example Code | Result/ Output |
|----------|-------------------------------------------------------------|--------------|----------------|
| \+       | Addition                                                    | x + y        | [1] 7          |
| \-       | Subtraction                                                 | x - y        | [1] -3         |
| \*       | Multiplication                                              | x \* y       | [1] 10         |
| /        | Division                                                    | x / y        | [1] 0.4        |
| %%       | Modulus (returns the remainder after division)              | y %% x       | [1] 1          |
| %/%      | Integer division (returns an integer value after division)  | y%/% x       | [1] 2          |
| \^       | Exponent                                                    | y \^ x       | [1]25          |

### **Relational operators**

| **Operator** | **Description**          | Example Code | Result/Output |
|--------------|--------------------------|--------------|---------------|
| \<           | Less than                | x \< y       | [1] TRUE      |
| \>           | Greater than             | x \> y       | [1] FALSE     |
| \<=          | Less than or equal to    | x \< = 2     | [1] TRUE      |
| \>=          | Greater than or equal to | y \>= 10     | [1] FALSE     |
| ==           | Equal to                 | y == 5       | [1] TRUE      |
| !=           | Not equal to             | x != 2       | [1] FALSE     |

### **Logical operators**

| **Operator** | **Description**          |
|--------------|--------------------------|
| &            | Element-wise logical AND |
| &&           | Logical AND              |
| \|           | Element-wise logical OR  |
| \|\|         | Logical OR               |
| !            | Logical NOT              |

```{r}
x <- c(3, 5, 7)

y <- c(2, 4, 6)


```

```{r}
x < 5 & y < 5 # all elements tested
```

```{r}
x < 5 && y < 5 # first element tested
```

## Organize your data

```{r}
penguins %>%
  arrange(-bill_length_mm)
```

#### Arrange

```{r}
penguins_2 <- penguins %>%
  arrange(-bill_length_mm) # Organized desc - form penguins

penguins_2 # bill is pico in spanish
```

```{r}
library(tidyverse)
```

#### Group By

```{r}

# Careful with drop_na
penguins %>% 
  group_by(island) %>%
  drop_na() %>% # eleminated na values in the data set
  summarize(mean_bill_length_mm = mean(bill_length_mm))

```

```{r}
penguins %>% 
  group_by(island) %>%
  drop_na() %>% # eleminated na values in the data set
  summarize(max_bill_length_mm = max(bill_length_mm))
```

```{r}
penguins %>% 
  drop_na() %>%
  summarize(min(bill_length_mm))
```

```{r}
penguins %>% 
  group_by(island) %>%
  drop_na() %>% # eleminated na values in the data set
  summarize(min_bill_length_mm = min(bill_length_mm))
```

```{r}
penguins %>% 
  group_by(species, island) %>%
  drop_na() %>% # eleminated na values in the data set
  summarize(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))
```

#### Filters

```{r}
penguins %>% 
  filter(species == 'Adelie') 

```

## Transforming data

```{r}
id <- c(1:10)

name <- c("John Mendes", "Rob Stewart", "Rachel Abrahamson", "Christy Hickman", "Johnson Harper", "Candace Miller", "Carlson Landy", "Pansy Jordan", "Darius Berry", "Claudia Garcia")

job_title <- c("Professional", "Programmer", "Management", "Clerical", "Developer", "Programmer", "Management", "Clerical", "Developer", "Programmer")

employee <- data.frame(id, name, job_title)
```

```{r}
print(employee)
```

#### Separate

```{r}
employee_name <- separate(employee, name, into=c("first_name", "last_name"), sep= ' ')
```

```{r}
print(employee_name)
```

#### Unite

```{r}
unite(employee_name, 'name', first_name, last_name) # By default separeted with _ 
```

```{r}
unite(employee_name, 'name', first_name, last_name, sep = ' ') # You can indicated
```

#### Mutate

```{r}
penguins_edit <- penguins %>% 
  mutate(body_mass_kg = body_mass_g/1000, flipper_length_m = flipper_length_mm/1000)
```

```{r}
View(penguins_edit)
```

## Wide to long with tidyr

**Wide data** has observations across several columns. Each column
contains data from a different condition of the variable. In this
example, different years. 

![screenshot of a spreadsheet in wide format. there are 7 rows shown and
8
columns](https://d3c33hcgiwev3.cloudfront.net/imageAssetProxy.v1/m9JdRN6aTQCSXUTemq0AVA_4909c4cc120d496c99895460c2ca27f4_Screenshot-2021-01-27-at-2.26.11-PM.png?expiry=1622764800000&hmac=g8HJZjnA-1dNtSKEIQkM5MvDcO45Q4TUiMjuShbLolY)

And, to review what you already learned about the difference, **long
data** has all the observations in a single column, and variables in
separate columns. 

![screenshot of spreadsheet in long format. there are 12 rows shown and
3
columns](https://d3c33hcgiwev3.cloudfront.net/imageAssetProxy.v1/AKr0lJTgSaCq9JSU4CmgtA_7608bba87aa04ab4b1695c3b3a22325a_Screenshot-2021-01-27-at-2.29.39-PM.png?expiry=1622764800000&hmac=B0jaSu6NdRwHqvOChnN6P6b3-W_-2jQfBqDIP3c06D0)

### The pivot_longer and pivot_wider functions

[`pivot_longer()`](https://tidyr.tidyverse.org/reference/pivot_longer.html)
can work with multiple value variables that may have different types,
inspired by the enhanced `melt()` and `dcast()` functions provided by
the [data.table](https://github.com/Rdatatable/data.table/wiki) package
by Matt Dowle and Arun Srinivasan.

[`pivot_longer()`](https://tidyr.tidyverse.org/reference/pivot_longer.html)
and
[`pivot_wider()`](https://tidyr.tidyverse.org/reference/pivot_wider.html)
can take a data frame that specifies precisely how metadata stored in
column names becomes data variables (and vice versa), inspired by the
[cdata](https://winvector.github.io/cdata/) package by John Mount and
Nina Zumel.

```{r}
library(tidyr)
library(dplyr)
library(readr)
```

```{r}
relig_income
```

This dataset contains three variables:

-   `religion`, stored in the rows,

-   `income` spread across the column names, and

-   `count` stored in the cell values.

    ```{r}
    relig_income %>% 
      pivot_longer(!religion, names_to = "income", values_to = "count")
    ```

```{r}
billboard

```

```{r}
billboard %>% 
  pivot_longer(
    cols = starts_with("wk"), 
    names_to = "week", 
    values_to = "rank",
    values_drop_na = TRUE
  )
```

```{r}
billboard %>% 
  pivot_longer(
    cols = starts_with("wk"), 
    names_to = "week", 
    names_prefix = "wk",
    names_transform = list(week = as.integer),
    values_to = "rank",
    values_drop_na = TRUE,
  )
```

```{r}
billboard %>% 
  pivot_longer(
    cols = starts_with("wk"), 
    names_to = "week", 
    names_transform = list(week = readr::parse_number),
    values_to = "rank",
    values_drop_na = TRUE,
  )

```

### Many variables in column names

-   The `new_`/`new` prefix indicates these are counts of new cases.
    This dataset only contains new cases, so we'll ignore it here
    because it's constant.

-   `sp`/`rel`/`ep` describe how the case was diagnosed.

-   `m`/`f` gives the gender.

-   `014`/`1524`/`2535`/`3544`/`4554`/`65` supplies the age range.

```{r}
who
```

```{r}
who %>% pivot_longer(
  cols = new_sp_m014:newrel_f65,
  names_to = c("diagnosis", "gender", "age"), 
  names_pattern = "new_?(.*)_(.)(.*)",
  values_to = "count"
)
```

```{r}
who %>% pivot_longer(
  cols = new_sp_m014:newrel_f65,
  names_to = c("diagnosis", "gender", "age"), 
  names_pattern = "new_?(.*)_(.)(.*)",
  names_transform = list(
    gender = ~ readr::parse_factor(.x, levels = c("f", "m")),
    age = ~ readr::parse_factor(
      .x,
      levels = c("014", "1524", "2534", "3544", "4554", "5564", "65"), 
      ordered = TRUE
    )
  ),
  values_to = "count",
)
```

## Wider

```{r}
fish_encounters
```

```{r}
fish_encounters %>% pivot_wider(names_from = station, values_from = seen)
```

```{r}
fish_encounters %>% pivot_wider(
  names_from = station, 
  values_from = seen,
  values_fill = 0
)
```

```{r}
warpbreaks <- warpbreaks %>% as_tibble() %>% select(wool, tension, breaks)
warpbreaks
```

```{r}
warpbreaks %>% count(wool, tension)
```

```{r}
warpbreaks %>% pivot_wider(names_from = wool, values_from = breaks)
```

```{r}
warpbreaks %>% 
  pivot_wider(
    names_from = wool, 
    values_from = breaks,
    values_fn = list(breaks = mean)
  )
```
