---
title: "R Notebook Wrangle"
output: html_notebook
---

## Data Cleaning with R

The process of identifying, correcting, or removing inaccurate raw data for downstream purposes. Or, more colloquially, an unglamorous yet wholely necessary first step towards an analysis-ready.

### Check List

-   [ ]  Familiarize yourself with the data set

-   [ ] Check for structural errors

-   [ ] Check for data irregularities

-   [ ] Decide how to deal with missing values

-   [ ] Document data versions and changes made

### Step 1: Familiarize yourself with the data set

An important "pre-data cleaning" Step is **Domain Knowledge**.

> Be sure to read up or ask an expert **before** diving into data cleaning if your variables don't make sense to you at first.sure to read up or ask an expert **before** diving into data cleaning if your variables don't make sense to you at first.

```{r}
file.info("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")$size
```

The data file we're utilizing is 1104203 bytes or 1.01MB.

```{r}
df = read.csv("./mental-heath-in-tech-2016_20161114.csv/mental-heath-in-tech-2016_20161114.csv")

```

```{r}
head(df)
```

```{r}
str(df) # Show structure of data frame
```

#### Simple Fast Exploratory Data Analysis in R with DataExplorer Package

```{r}
#Install if the package doesn't exist 
update.packages()

```

```{r}
install.packages('DataExplorer')
library(DataExplorer)
```
