---
title: "R Notebook"
output: html_notebook
---

## Dates and times in R

```{r}
install.packages("tidyverse") # Una libreria impresindible

```

```{r}
library(tidyverse)
```

```{r}
library(lubridate)
```

### Types

-   A date `("2016-08-16")`

-   A time within a day `(“20:11:59 UTC")`

-   And a date-time. This is a date plus a time `("2018-03-31 18:15:48 UTC")`

```{r}
today() # Nos muestra la fecha actual

```

```{r}
now() # Nos devuelve la fecha y la hora actual
```

## Trabajando con R

Al usar R tenemos tres formas de crear fechas-tiempo

-   Desde un String

-   Desde una fecha individual

-   Desde un existente date/time object

El estandar que usa R para crear fechas es `yyyy-mm-dd`

### **Converting from strings** 

```{r}
ymd("2021-01-20")
```

```{r}
mdy("January 20th, 2021")
```

```{r}
dmy("20-Jan-2021")
```

```{r}
ymd(20210120)

```

### **Creating date-time components**

```{r}
ymd_hms("2021-01-20 20:11:59")
```

```{r}
mdy_hm("01/20/2021 08:01")
```

```{r}
as_date(now()) #to convert a date-time to a date
```

# Data structures

### **Data frames**

Data frames are the most common way of storing and analyzing data in R,
so it's important to understand what they are and how to create them.

There are a few key things to keep in mind when you are working with data frames: 

-   First, columns should be named. 

-   Second, data frames can include many different types of data, like numeric, logical, or character.

-   Finally, elements in the same column should be of the same type.

If you need to manually create a data frame in R, you can use the **data.frame()** function.

```{r}
data.frame(x = c(1,2,3), y = c(1.5, 5.5, 7.5))
```

## Files

Use the **dir.create** function to create a new folder, or
directory, to hold your files. Place the name of the folder in the
parentheses of the function.

```{r}
file.create ("new_text_file.txt")
```

```{r}
file.copy ("new_text_file.txt" , "./CC-RBasics")
```

```{r}
# Borrar file
unlink ("new_text_file.txt")

```

### Matrices (optional)

A **matrix** is a two-dimensional collection of data
elements. This means it has both rows and columns. By contrast, a vector
is a one-dimensional sequence of data elements. But like vectors,
matrices can only contain a single data type. For example, you can't
have both logicals and numerics in a matrix. 

```{r}

matrix(c(3:8), nrow = 2)
```

```{r}
matrix(c(3:8), ncol = 2)

```
